---
title: "ETC5521 Assignment 1"
subtitle: "Your title"
team: Magpie
author:
  - Junhao Wang 
  - Priya Ravindra Dingorkar
date: "`r Sys.Date()`"
output: 
  html_document
---

[This assignment is for ETC5521 Assignment 1 by Team `r rmarkdown::metadata$team` comprising of `r knitr::combine_words(rmarkdown::metadata$author)`.]{style="color:#006DAE;"}



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(kableExtra)
```

# Introduction and motivation

[FILL] Give the bigger picture of the data, and inspire the reader to learn more about the problem by reading your analysis. 

- Junhao, add the introduction and motivation


```{r dat-load, message = FALSE, echo=FALSE}
tuition_cost <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_cost.csv')

tuition_income <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_income.csv') 

salary_potential <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/salary_potential.csv')

historical_tuition <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/historical_tuition.csv')

diversity_school <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/diversity_school.csv')
```

# Research Question

Primary Question


Secondary Questions


# Data description

It is through data collection that a business or management has the quality information that they need for further analysis, study, and research to make informed decisions. We will perform exploratory, data review in this study on the subject of [**College Tuition, Diversity, and Pay Dataset**](https:/github.com / rfordatascience / tidytuesday / blob / master / data/2020/2020-03-10 / readme.md). This collection of data primarily receives all the information from various sources but originally from the US Department of Education.

The data, taken from the original pages, turned out to be very large in size, which is why the author, for the convenience of the audience and to produce informative results, filtered the data into five comma-separated documents using web scraping in R with the support of the rvest package.

Let's look at the small snippets of the various data sets, and get a better understanding of the variables

1. The diversity_school data set comes [The Chronicle of Higher Education](https://www.chronicle.com/article/student-diversity-at-4-725-institutions).

```{r ds, echo=FALSE}
head(diversity_school) %>% kableExtra::kable() %>% kableExtra::kable_styling(c("hover" , "striped"))
```

- Name has data type character properties that describe the school name
- Total Enrollment has double datatype properties representing the total number of students enrolled
- State with character datatype representing the various states
- Category with character properties defining the group Group / Racial / Gender
- Enrollment with double datatype properties specifying the specific enrollment.

```{r, message=FALSE, echo=TRUE}
glimpse(diversity_school)
```

2. The historical_tution dataset been obtained from [U.S. Department of Education, National Center for Education Statistics](https://nces.ed.gov/fastfacts/display.asp?id=76)

```{r ht, echo=FALSE}
head(historical_tuition) %>% kableExtra::kable() %>% kableExtra::kable_styling(c("hover" , "striped"))
```

- type variable having character datatype describes the Type of school (All, Public, Private)
- year having character format describes the corresponding Academic year
- tuition_type having	character datatype refers to duration of the degree, Tuition Type All Constant (dollar inflation adjusted), 4 year degree constant, 2 year constant, Current to year, 4 year current, 2 year current
- tuition_cost having	double datatype represents the education fees/Tuition cost in USD

```{r, message=FALSE, echo=TRUE}
glimpse(historical_tuition)
```


# Analysis and findings

[FILL] Should include at least one plot or numerical summary for each of your questions, that helps the reader arrive at an answer. You should also write paragraphs describing the methods, summaries and findings. 

# References
